---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# DisaggregateTS

<!-- badges: start -->
[![CRAN](https://www.r-pkg.org/badges/version/DisaggregateTS)](https://cran.r-project.org/package=DisaggregateTS)
<!-- badges: end -->

The goal of DisaggregateTS is to provide tools for temporal disaggregation, including: (1) high-dimensional and low-dimensional series generation for simulation studies; (2) a toolkit for temporal disaggregation and benchmarking using low-dimensional indicator series as proposed by Dagum and Cholette; (3) novel techniques by Mosley, Gibberd, and Eckley (2022) for disaggregating low-frequency series in the presence of high-dimensional indicator matrices.

## Installation

To install or update from [CRAN](https://cran.r-project.org/web/packages/DisaggregateTS/index.html)

``` r
# install.packages("DisaggregateTS")
install.packages("DisaggregateTS")
```

You can install the development version of DisaggregateTS from [GitHub](https://github.com/kavehsn/DisaggregateTS) with:

``` r
# install.packages("remotes")
remotes::install_github("kavehsn/DisaggregateTS")
```

## Example

The `TempDisaggDGP()` generates a high-frequency response vector y, following the relationship $y = X\beta+\epsilon$, where $X$ is a matrix of indicator series and $\beta$ is a potentially sparse coefficient vector. The low-frequency vector $Y$ is generated by aggregating $y$ according to a specified aggregation method. On the other hand, the `disaggregate()` function contains the traditional standard-dimensional temporal disaggregation methods proposed by Denton (1971), Dagum and Cholette (2006), Chow and Lin (1971), Fernández (1981) and Litterman (1983), and the high-dimensional methods of Mosley et al. (2022). 

```{r example}
library(DisaggregateTS)
## basic example code
data <- TempDisaggDGP(n_l=25, n=100, p=10, rho=0.5) 
X <- data$X_Gen
Y <- data$Y_Gen 
fit_chowlin <- disaggregate(Y=Y,X=X,method='Chow-Lin') 
y_hat = fit_chowlin$y_Est
```

Please report any bugs to [Kaveh Nobari](mailto:k.salehzadeh-nobari@imperial.ac.uk).
